---
title: |
  STA 209: Introduction to Time Series Analysis
subtitle: |
  Homework 2
author: |
  Rento Saijo
date: 'February 19, 2026'
fontsize: 10pt
documentclass: extarticle
geometry: top=1.5in, bottom=1.5in, left=1.5in, right=1.5in
urlcolor: blue
linkcolor: blue
citecolor: blue
output:
  bookdown::pdf_document2:
    keep_tex: true
    toc: false
    fig_width: 5
    fig_height: 3.5
    fig_caption: true

header-includes:
  - |
    \usepackage{xcolor}
    \usepackage{hyperref}
    \hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue, citecolor=blue}
    \usepackage{float}
    \usepackage{fvextra}
    \usepackage{xcolor}
    \usepackage{fancyhdr}
    \usepackage{lastpage}
    \usepackage{soul}
    \usepackage{etoolbox}
    \usepackage{microtype}
    \usepackage{tcolorbox}
    \usepackage{enumitem}
    \usepackage{fancyvrb}
    \usepackage{xspace}
    \allowdisplaybreaks
    \setcounter{tocdepth}{2}
    \setcounter{secnumdepth}{0}
    \floatplacement{figure}{H}
    \makeatletter
    \newcommand{\@subtitle}{}
    \newcommand{\subtitle}[1]{\gdef\@subtitle{#1}}
    \newcommand{\DocSubtitle}{\@subtitle}
    \renewcommand{\maketitle}{
      \thispagestyle{plain}
      \null
      \vfill
      \begin{center}
        {\Large \textsc{\@title} \par}\vskip 0.5em
        {\LARGE \bfseries \@subtitle \par}\vskip 0.75em
        {\large \@author \par}\vskip 0.5em
        {\normalsize \@date \par}
      \end{center}
      \vfill
      \tableofcontents
      \vspace{1em}
      \clearpage
    }
    \AfterEndEnvironment{Highlighting}{\setcounter{CodeLine}{\numexpr\value{CodeLine}+\FV@CodeLineNo\relax}}
    \makeatother
    \fancypagestyle{plain}{
      \fancyhf{}
      \renewcommand{\headrulewidth}{0pt}
      \renewcommand{\footrulewidth}{0pt}
    }
    \pagestyle{fancy}
    \fancyhf{}
    \fancyfoot[C]{\small\textsc{Page}~\thepage~\textsc{of}~\pageref{LastPage}}
    \fancyhead[L]{\textsc{Rento Saijo}}
    \fancyhead[R]{\textsc{\DocSubtitle}}
    \fancyhead[C]{\textsc{\rightmark}}
    \renewcommand{\headrulewidth}{0.5pt}
    \renewcommand{\footrulewidth}{0.5pt}
    \newcounter{CodeLine}
    \newcounter{cell}
    \AtBeginEnvironment{Highlighting}{\stepcounter{cell}}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{
      frame=single,
      rulecolor=\color{black},
      framesep=2mm,
      label=\footnotesize\textbf{Cell \thecell},
      labelposition=topline,
      commandchars=\\\{\},
      breaklines=true,
      breakanywhere=false,
      breaksymbol={},
      numbers=left,
      numbersep=3pt,
      firstnumber=\numexpr\value{CodeLine}+1\relax,
      fontsize=\small
    }
    \DefineVerbatimEnvironment{verbatim}{Verbatim}{
      breaklines=true,
      breakanywhere=true,
      numbers=left,
      numbersep=3pt,
      firstnumber=1,
      fontsize=\footnotesize
    }
    \DefineVerbatimEnvironment{snippet}{Verbatim}{
      breaklines=true,
      breakanywhere=true,
      fontsize=\footnotesize,
      frame=single,
      listparameters=\setlength{\topsep}{\baselineskip}\setlength{\partopsep}{0pt}
    }
    \newtcolorbox{problem}[1][]{
      title={\textsc{#1}}
    }
    \newenvironment{alphenum}{
      \begin{enumerate}[
        label=(\alph*),
        itemsep=3pt,
        parsep=0pt,
        topsep=6pt
      ]
    }{
      \end{enumerate}
    }
    \newenvironment{items}{
      \begin{itemize}[
        itemsep=3pt,
        parsep=0pt,
        topsep=6pt
      ]
    }{
      \end{itemize}
    }
    \renewcommand{\contentsname}{Table of Contents}
    \newcommand{\E}{\mathrm{E}}
    \newcommand{\Var}{\mathrm{Var}}
    \newcommand{\R}{\textsf{R}\xspace}
    \newcommand{\Pois}{\mathrm{Pois}}
    \newcommand{\SD}{\mathrm{SD}}
    \newcommand{\SE}{\mathrm{SE}}
---

```{r setup, include = FALSE}
# ----- Setup ----- #

# Set options.
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = 'center')
options(width = 125)
options(scipen = 999)

# Load libraries.
suppressMessages(library(tidyverse))
suppressMessages(library(readxl))
suppressMessages(library(fs))
suppressMessages(library(plotly))
suppressMessages(library(htmlwidgets))
suppressMessages(library(astsa))
suppressMessages(library(nhlscraper))
suppressMessages(library(showtext))
suppressMessages(library(Cairo))

# Set fonts.
sysfonts::font_add(
  family  = 'cmuserif',
  regular = '/Users/rsai_91/Library/Fonts/cmu.serif-roman.ttf',
  bold    = '/Users/rsai_91/Library/Fonts/cmu.serif-bold.ttf',
  italic  = '/Users/rsai_91/Library/Fonts/cmu.serif-italic.ttf',
  bolditalic = '/Users/rsai_91/Library/Fonts/cmu.serif-bolditalic.ttf'
)
showtext::showtext_opts(dpi = 300)
showtext::showtext_auto(enable = TRUE)
knitr::opts_chunk$set(dev = 'CairoPDF')
par(family = 'cmuserif')
ggplot2::theme_set(ggplot2::theme_minimal(base_family = 'cmuserif', base_size = 10))

# Set seed.
set.seed(20060527)

# ----- Helpers ----- #

# Print LaTeX table from data.frame.
table_latex = function(data, col_names = NULL, caption = NULL) {
  data %>%
    knitr::kable(
      format    = 'latex',
      booktabs  = TRUE,
      escape    = FALSE,
      col.names = col_names,
      caption   = caption,
      linesep   = '',
      align     = rep('c', ncol(data)),
      ) %>%
    kableExtra::kable_styling(
      position = 'center',
      latex_options = c('HOLD_position'))
}

# Plot with LaTeX.
with_par <- function(expr) { 
    op <- par(no.readonly = TRUE)
    on.exit(par(op))
    force(expr) 
}
with_family <- function(expr, family = 'cmuserif') {
    op <- par(no.readonly = TRUE)
    on.exit(par(op))
    par(family = family)
    force(expr) 
}

# Read from Excel.
ts_seasonal <- function(filename, nums = 1:4) {
  data.seasonal <- matrix(0, nrow = 1, ncol = 5)
  for(i in 1:length(nums)){
    dataexcel <- suppressMessages(readxl::read_excel(filename, sheet = nums[i], range = 'D7:H77'))
    dataexcel <- dataexcel[-1,-2]
    dataexcel <- dplyr::filter(dataexcel, !is.na(dataexcel$Avg))
    data <- dplyr::mutate(dataexcel, Month = nums[i])
    colnames(data.seasonal) <- colnames(data)
    data.seasonal <- rbind(data.seasonal, data)
  }
  data.seasonal <- data.seasonal[-1, ]
  data.seasonal <- dplyr::mutate(data.seasonal, Year = readr::parse_number(as.character(Year)))
  data.seasonal <- dplyr::filter(data.seasonal, !is.na(Year))
  data.seasonal <- dplyr::arrange(data.seasonal, Year)
  dplyr::mutate(data.seasonal, Month = factor(Month, levels = 1:4, labels = c('January', 'April', 'July', 'October')))
}

# Make ts objects (freq = 4) for Min/Avg/Max, starting at the earliest year.
make_site_ts <- function(df) {
  df <- dplyr::mutate(df, Year = readr::parse_number(as.character(Year)))
  df <- dplyr::filter(df, !is.na(Year))
  df$Season <- as.numeric(df$Month)
  df <- dplyr::arrange(df, Year, Season)
  start_year <- min(df$Year, na.rm = TRUE)
  ts_min <- stats::ts(df$Min, start = c(start_year, 1), frequency = 4)
  ts_avg <- stats::ts(df$Avg, start = c(start_year, 1), frequency = 4)
  ts_max <- stats::ts(df$Max, start = c(start_year, 1), frequency = 4)
  list(ts_min = ts_min, ts_avg = ts_avg, ts_max = ts_max, start_year = start_year)
}
```

# Disclosure

GPT-5.3-Codex was used to create the `YAML` portion and some `LaTeX` code to format the text/equations nicely. Page formatting code was also provided by Derin Gezgin. In the setup chunk, libraries were loaded and some helper functions were defined including but not limited to `table_latex()` and `with_family()`. See the original `RMD` file [here](https://github.com/RentoSaijo/STA209/blob/main/Homework2.Rmd) for more details.

\newpage

# Problem 1

\begin{problem}[Problem 1]
Consider the Southern oscillation index data (\texttt{soi}) from Homework 1. Use this data to do the following in R and report your results. Use periodicity \(s = 12\). Fit the following five models:
\begin{align*}
\textbf{Model I:}\quad
\text{SOI}_t
&= \beta_0 + \beta_1 t + W_t,\\
\textbf{Model II:}\quad
\text{SOI}_t
&= \beta_0 + \beta_1 t + \frac{\beta_2 t^2}{2!} + W_t,\\
\textbf{Model III:}\quad
\text{SOI}_t
&= \beta_0 + \alpha_1\cos\!\left(\frac{2\pi t}{s}\right) + \alpha_2\sin\!\left(\frac{2\pi t}{s}\right) + W_t,\\
\textbf{Model IV:}\quad
\text{SOI}_t
&= \beta_0 + \beta_1 t + \alpha_1\cos\!\left(\frac{2\pi t}{s}\right) + \alpha_2\sin\!\left(\frac{2\pi t}{s}\right) + W_t,\\
\textbf{Model V:}\quad
\text{SOI}_t
&= \beta_0 + \beta_1 t
+ \alpha_1\cos\!\left(\frac{2\pi t}{s}\right)
+ \alpha_2\sin\!\left(\frac{2\pi t}{s}\right)\\
&\quad + \alpha_3\cos\!\left(\frac{4\pi t}{s}\right)
+ \alpha_4\sin\!\left(\frac{4\pi t}{s}\right)
+ W_t.
\end{align*}
\end{problem}

The Southern Oscillation Index (SOI) dataset is a monthly time series with 453 observations covering the years 1950 through 1987. It is represented in the form \texttt{Time-Series [1:453]} (notated as running from 1950 to 1988 in the printed format) and is shown as a numeric sequence (e.g., \texttt{0.377, 0.246, 0.311, 0.104, -0.016, 0.235, \dots}). The series is intended to pair with \texttt{rec} (Recruitment) for related analyses. The data were furnished by Dr.\ Roy Mendelssohn of the Pacific Fisheries Environmental Laboratory, NOAA (personal communication). Demonstrations of \texttt{astsa} capabilities are referenced under \emph{FUN WITH ASTSA}, and the most recent version of the package is maintained on GitHub along with the \texttt{NEWS} and \texttt{ChangeLog} files; additional webpages for the associated texts and help using \textsf{R} for time series analysis are available at the author’s site.

\newpage

## Problem 1 (a)

\begin{problem}[Problem 1 (a)]
Report the five fitted models.
\end{problem}

```{r}
# Load SOI data and define periodicity.
data(soi)
s <- 12
Y <- soi
t <- time(Y)
ts2 <- t^2 / factorial(2)
cs1 <- cos(2 * pi * t / s)
sn1 <- sin(2 * pi * t / s)
cs2 <- cos(4 * pi * t / s)
sn2 <- sin(4 * pi * t / s)

# Fit all five models.
m1 <- lm(Y ~ t)
m2 <- lm(Y ~ t + ts2)
m3 <- lm(Y ~ cs1 + sn1)
m4 <- lm(Y ~ t + cs1 + sn1)
m5 <- lm(Y ~ t + cs1 + sn1 + cs2 + sn2)

# See model summaries.
summary(m1)
summary(m2)
summary(m3)
summary(m4)
summary(m5)
```

The fitted models are:
\begin{align*}
\widehat{\textbf{Model I:}}\quad
\widehat{\text{SOI}}_t
&= 13.704 - 0.007\,t,\\[4pt]
\widehat{\textbf{Model II:}}\quad
\widehat{\text{SOI}}_t
&= -496.761 + 0.512\,t - 0.000\,\frac{t^2}{2!},\\[4pt]
\widehat{\textbf{Model III:}}\quad
\widehat{\text{SOI}}_t
&= 0.079 - 0.001\,\cos\!\left(\frac{2\pi t}{12}\right) - 0.031\,\sin\!\left(\frac{2\pi t}{12}\right),\\[4pt]
\widehat{\textbf{Model IV:}}\quad
\widehat{\text{SOI}}_t
&= 13.576 - 0.007\,t - 0.011\,\cos\!\left(\frac{2\pi t}{12}\right) - 0.010\,\sin\!\left(\frac{2\pi t}{12}\right),\\[4pt]
\widehat{\textbf{Model V:}}\quad
\widehat{\text{SOI}}_t
&= 12.797 - 0.006\,t - 0.008\,\cos\!\left(\frac{2\pi t}{12}\right) - 0.009\,\sin\!\left(\frac{2\pi t}{12}\right)\\
&\quad - 0.043\,\cos\!\left(\frac{4\pi t}{12}\right) + 0.081\,\sin\!\left(\frac{4\pi t}{12}\right),
\end{align*}
where \(\text{SOI}_t\) denotes the Southern Oscillation Index at time \(t\).

\newpage

## Problem 1 (b)

\begin{problem}[Problem 1 (b)]
Prepare a summary table using number of regression parameters, \(R_a^2\), AIC, AICC and BIC.
\end{problem}

```{r}
# Build table of fit criteria.
fit_table <- function(fit) {
  n   <- length(resid(fit))
  r   <- length(coef(fit))
  sse <- sum(resid(fit)^2)
  data.frame(
    Parameters = r,
    R2a  = summary(fit)$adj.r.squared,
    AIC  = AIC(fit) / n,
    AICC = log(sse / n) + ((n + r) / (n - r - 2)),
    BIC  = AIC(fit, k = log(n)) / n
  )
}
comp1 <- dplyr::bind_rows(
  'Model I'   = fit_table(m1),
  'Model II'  = fit_table(m2),
  'Model III' = fit_table(m3),
  'Model IV'  = fit_table(m4),
  'Model V'   = fit_table(m5),
  .id = 'Model'
)
comp1$Parameters <- as.integer(comp1$Parameters)
comp1 <- comp1 %>%
  dplyr::mutate(
    R2a  = round(R2a, 3),
    AIC  = round(AIC, 3),
    AICC = round(AICC, 3),
    BIC  = round(BIC, 3)
  )
table_latex(comp1, col_names = c('Model', '\\# Parameters', '$R_a^2$', 'AIC', 'AICC', 'BIC'))
```

\newpage

## Problem 1 (c)

\begin{problem}[Problem 1 (c)]
Using the summary in (b), which model is selected to explain SOI and why?
\end{problem}

For explanation, we prioritize adjusted \(R^2\). Model V has the largest adjusted \(R^2\), with \(R_a^2 = 0.058\). Also, its overall F-test is significant with \(F = 6.520\) and \(p = 0.000\) (although all others have fairly low values as well). All in all, though, \textbf{Model V} is selected for explaining SOI variation among these candidates.

\newpage

## Problem 1 (d)

\begin{problem}[Problem 1 (d)]
Based on AIC, AICC and BIC, which model is selected and what is it best for?
\end{problem}

AIC and AICC are minimized by \textbf{Model V}, with \(AIC = 0.875\) and \(AICC = -0.962\), but BIC is minimized by \textbf{Model I}, with \(BIC = 0.916\). Since these criteria disagree and BIC applies the strongest complexity penalty, we select \textbf{Model I}. This selected model maximizes the probability that the sample data are from the suggested model.

\newpage

## Problem 1 (e)

\begin{problem}[Problem 1 (e)]
Remove the last 30 values and compare future forecast error across models using ME, MPE, MSE, MAE, and MAPE. Include these with \(R_a^2\), AIC, AICC and BIC in a table. Does the selected model match part (c)?
\end{problem}

```{r}
# Hold out the last 30 observations.
h     <- 30
n_all <- length(Y)
cut   <- n_all - h
Y_train   <- stats::window(Y, end = time(Y)[cut])
Y_test    <- stats::window(Y, start = time(Y)[cut + 1])
t_train   <- time(Y_train)
t_test    <- time(Y_test)
ts2_train <- t_train^2 / factorial(2)
ts2_test  <- t_test^2 / factorial(2)
cs1_train <- cos(2 * pi * t_train / s)
sn1_train <- sin(2 * pi * t_train / s)
cs2_train <- cos(4 * pi * t_train / s)
sn2_train <- sin(4 * pi * t_train / s)
cs1_test  <- cos(2 * pi * t_test / s)
sn1_test  <- sin(2 * pi * t_test / s)
cs2_test  <- cos(4 * pi * t_test / s)
sn2_test  <- sin(4 * pi * t_test / s)

# Refit on training data.
f1 <- lm(Y_train ~ t_train)
f2 <- lm(Y_train ~ t_train + ts2_train)
f3 <- lm(Y_train ~ cs1_train + sn1_train)
f4 <- lm(Y_train ~ t_train + cs1_train + sn1_train)
f5 <- lm(Y_train ~ t_train + cs1_train + sn1_train + cs2_train + sn2_train)

# Forecast the held-out 30 observations.
p1 <- predict(f1, newdata = data.frame(t_train = t_test))
p2 <- predict(f2, newdata = data.frame(t_train = t_test, ts2_train = ts2_test))
p3 <- predict(f3, newdata = data.frame(cs1_train = cs1_test, sn1_train = sn1_test))
p4 <- predict(f4, newdata = data.frame(t_train = t_test, cs1_train = cs1_test, sn1_train = sn1_test))
p5 <- predict(f5, newdata = data.frame(t_train = t_test, cs1_train = cs1_test, sn1_train = sn1_test, cs2_train = cs2_test, sn2_train = sn2_test))

# Forecast error metrics (Lecture 5 style).
fcst_metrics <- function(y_true, y_hat) {
  e <- y_true - y_hat
  data.frame(
    ME = mean(e),
    MPE = 100 * mean(e / y_true),
    MSE = mean(e^2),
    MAE = mean(abs(e)),
    MAPE = 100 * mean(abs(e / y_true))
  )
}
fc_table <- dplyr::bind_rows(
  'Model I' = fcst_metrics(Y_test, p1),
  'Model II' = fcst_metrics(Y_test, p2),
  'Model III' = fcst_metrics(Y_test, p3),
  'Model IV' = fcst_metrics(Y_test, p4),
  'Model V' = fcst_metrics(Y_test, p5),
  .id = 'Model'
)

# Training-fit criteria for the same refit models.
comp_train <- dplyr::bind_rows(
  'Model I' = fit_table(f1),
  'Model II' = fit_table(f2),
  'Model III' = fit_table(f3),
  'Model IV' = fit_table(f4),
  'Model V' = fit_table(f5),
  .id = 'Model'
)

# Round fit criteria table.
comp_fit <- comp_train %>%
  dplyr::mutate(
    Parameters = as.integer(Parameters),
    R2a = round(R2a, 3),
    AIC = round(AIC, 3),
    AICC = round(AICC, 3),
    BIC = round(BIC, 3)
  )

# Round forecast error table.
comp_fc <- fc_table %>%
  dplyr::mutate(
    ME = round(ME, 3),
    MPE = round(MPE, 3),
    MSE = round(MSE, 3),
    MAE = round(MAE, 3),
    MAPE = round(MAPE, 3)
  )

# Print two tables.
table_latex(
  comp_fit,
  col_names = c('Model', '\\# Parameters', '$R_a^2$', 'AIC', 'AICC', 'BIC')
)
table_latex(
  comp_fc,
  col_names = c('Model', 'ME', 'MPE', 'MSE', 'MAE', 'MAPE')
)
```

Using future-forecast error, \textbf{Model V} is preferred by MSE, with \(MSE = 0.106\). MPE, MAE, and MAPE are slightly smaller for Model III, so the error criteria are not fully unanimous; however, with MSE as the main forecasting criterion, Model V is selected. Therefore, the best model in part (e) matches part (c), but does not match part (d). I read a little bit more online on why MSE is preferred over MAE: my original thought was because it penalizes the further errors more, but that doesn't really seem to be the main concern; it said something about differentiability, but I couldn't necessarily understand why calculus is involves. Hopefully, I get to learn in Mathematical Statistics. If you have a quick intuitive explanation, would love to know!

\newpage

# Problem 2

\begin{problem}[Problem 2]
For the New Haven Average Temperature time series, fit the following trend and seasonality models:
\begin{align*}
\textbf{Model I:}\quad
\text{AvgTemp}_t
&= \beta_0 + \beta_1 t
+ \alpha_1 \cos\!\left(\frac{2\pi t}{s}\right) + \alpha_2 \sin\!\left(\frac{2\pi t}{s}\right) + W_t, \\
\textbf{Model II:}\quad
\text{AvgTemp}_t
&= \beta_0 + \beta_1 t
+ \frac{\beta_2 t^2}{2!}
+ \alpha_1 \cos\!\left(\frac{2\pi t}{s}\right) + \alpha_2 \sin\!\left(\frac{2\pi t}{s}\right) + W_t, \\
\textbf{Model III:}\quad
\text{AvgTemp}_t
&= \beta_0 + \beta_1 t
+ \frac{\beta_2 t^2}{2!}
+ \alpha_1 \cos\!\left(\frac{2\pi t}{s}\right)
+ \alpha_2 \sin\!\left(\frac{2\pi t}{s}\right) \\
&\quad + \alpha_3 \cos\!\left(\frac{4\pi t}{s}\right)
+ \alpha_4 \sin\!\left(\frac{4\pi t}{s}\right)
+ W_t.
\end{align*}
\end{problem}

\newpage

## Problem 2 (a)

\begin{problem}[Problem 2 (a)]
Determine \(s\) using periodogram.
\end{problem}

```{r}
# Load New Haven average temperature series.
NewHaven <- ts_seasonal('data/Climate_Northeast_NewHavenCT.xlsx')
Y <- make_site_ts(NewHaven)$ts_avg

# Periodogram
x <- Y - mean(Y)
n <- length(x)
I <- abs(fft(x))^2 / n
half <- floor((n / 2) - 1)
f <- 0:half / n
P <- (4 / n) * I[1:(n / 2)]
index <- which(P == max(P))
s <- round(1 / f[index])

# Report peak frequency and period.
f[index]
s

# Plot periodogram.
with_family({
  plot(f, P, type = 'l',
       main = 'Periodogram for Avg. Temp. in New Haven',
       xlab = 'Frequency', ylab = 'P(frequency)', col = 'blue')
  abline(v = f[index], col = 'red', lty = 2, lwd = 2)
})
```

The periodogram has its dominant peak at frequency \(f = 0.250\), so
\[
s = \mathrm{round}\!\left(\frac{1}{0.250}\right) = 4.
\]
Therefore, the New Haven average temperature series is modeled with quarterly seasonality (\(s = 4\)).

\newpage

## Problem 2 (b)

\begin{problem}[Problem 2 (b)]
Report the three fitted models.
\end{problem}

```{r}
# Build regressors and fit the three models.
t    <- as.numeric(time(Y))
t2   <- t^2 / factorial(2)
cs1  <- cos(2 * pi * t / s)
sn1  <- sin(2 * pi * t / s)
cs2  <- cos(4 * pi * t / s)
sn2  <- sin(4 * pi * t / s)
fit1 <- lm(Y ~ t + cs1 + sn1)
fit2 <- lm(Y ~ t + t2 + cs1 + sn1)
fit3 <- lm(Y ~ t + t2 + cs1 + sn1 + cs2 + sn2)

# See model summaries.
summary(fit1)
summary(fit2)
summary(fit3)
```

The fitted models are:
\begin{align*}
\widehat{\textbf{Model I:}}\quad
\widehat{\text{AvgTemp}}_t
&= 4.233 + 0.004\,t - 0.045\,\cos\!\left(\frac{2\pi t}{4}\right) - 0.096\,\sin\!\left(\frac{2\pi t}{4}\right),\\[6pt]
\widehat{\textbf{Model II:}}\quad
\widehat{\text{AvgTemp}}_t
&= 5952.734 - 5.998\,t + 0.003\,\frac{t^2}{2!} - 0.019\,\cos\!\left(\frac{2\pi t}{4}\right) - 0.078\,\sin\!\left(\frac{2\pi t}{4}\right),\\[6pt]
\widehat{\textbf{Model III:}}\quad
\widehat{\text{AvgTemp}}_t
&= 5914.417 - 5.960\,t + 0.003\,\frac{t^2}{2!} - 0.017\,\cos\!\left(\frac{2\pi t}{4}\right) - 0.077\,\sin\!\left(\frac{2\pi t}{4}\right)\\
&\quad + 0.034\,\cos\!\left(\frac{4\pi t}{4}\right) + 0.137\,\sin\!\left(\frac{4\pi t}{4}\right),
\end{align*}
where \(\text{AvgTemp}_t\) denotes the New Haven average temperature at time \(t\).

\newpage

## Problem 2 (c)

\begin{problem}[Problem 2 (c)]
Prepare a summary table with number of regression parameters, \(R_a^2\), AIC, AICC and BIC. Which model best explains temperature changes and why?
\end{problem}

```{r}
# Collect model-fit criteria.
fit_table <- function(fit) {
  n   <- length(resid(fit))
  r   <- length(coef(fit))
  sse <- sum(resid(fit)^2)
  data.frame(
    Parameters = r,
    R2a  = summary(fit)$adj.r.squared,
    AIC  = AIC(fit) / n,
    AICC = log(sse / n) + ((n + r) / (n - r - 2)),
    BIC  = AIC(fit, k = log(n)) / n
  )
}
comp <- dplyr::bind_rows(
  'Model I'   = fit_table(fit1),
  'Model II'  = fit_table(fit2),
  'Model III' = fit_table(fit3),
  .id = 'Model'
)
comp$Parameters <- as.integer(comp$Parameters)
comp <- comp %>%
  dplyr::mutate(
    R2a  = round(R2a, 3),
    AIC  = round(AIC, 3),
    AICC = round(AICC, 3),
    BIC  = round(BIC, 3)
  )
table_latex(comp, col_names = c('Model', '\\# Parameters', '$R_a^2$', 'AIC', 'AICC', 'BIC'))
```

From the comparison table:
\begin{items}
\item All three adjusted \(R^2\) values are near zero (slightly negative), so none of these deterministic trend/seasonality models explains much variation in this New Haven quarterly-average series (we can see all of the regular \(R^2\) values are near zero as well, so that's reasonable, I think). Model I has the largest \(R_a^2\) (least negative) among the three.
\item Model I also has the smallest AIC, AICC, and BIC.
\end{items}

Therefore, \textbf{Model I} is selected: adding the quadratic and higher-harmonic terms does not improve fit enough to offset the complexity penalty, although, again, it's not like Model I is great either.
